---
title: "ISYE6644 Project Nutwell Summer2022"
author: "Brian Nutwell"
date: '2022-07-22'
output: html_document
---

# Initial setup module

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:\\Users\\bnutw\\OneDrive\\Documents\\GTMS Analytics\\ISYE6501\\SimPy files')
rm(list=ls())

library(plotly)       # bar and line charts
library(dplyr)
library(pheatmap)     # pretty heatmaps
library(RColorBrewer) # color palettes for heatmaps and plotly
library(FrF2)         # fractional factorial design for DOE
library(stats)        # basic stats including standard deviation

```

All the actual game functions are now running in Python.
Using R for the visualizations.

# Import data from raw output files for early sample games

```{r}

demogamedata_2 <- as.data.frame(read.csv(file='TavernGameHist-Jul8a.csv',header=TRUE,sep=",",na.strings=c("")))
demogamedata_1 <- as.data.frame(read.csv(file='TavernGameHist-Jul8b.csv',header=TRUE,sep=",",na.strings=c("")))
demo_3df <- as.data.frame(read.csv(file='TavernGameSampleRolls.csv',header=TRUE,sep=",",na.strings=c("")))

samplerunstowin <- as.data.frame(read.csv(file='TavernRunsToWinDist.csv',header=TRUE,sep=",",na.strings=c("")))
colnames(samplerunstowin) <- c('RunsToWin','Occurrence')

samplewinners <- as.data.frame(read.csv(file='TavernWinnersDist.csv',header=TRUE,sep=",",na.strings=c("None")))
colnames(samplewinners) <- c('Player','Wins')

```

# Histogram charts for dice roll occurrences to demonstrate they are following expected distributions

```{r}

roll_list <- c(demogamedata_1['LowBone'], demogamedata_1['HighBone'], demogamedata_2['LowBone'], 
    demogamedata_2['HighBone'])

c1df <- as.data.frame(table(unlist(roll_list)))
colnames(c1df) <- c('Die Roll','Count')


roll_list2 = list()
for (i in seq(1,4)) {
  coltoget <- paste('P',i,'_Roll',sep='')
  roll_list2 <- c(roll_list2,demogamedata_1[coltoget],demogamedata_2[coltoget])
}

c2df <- as.data.frame(table(unlist(roll_list2)))
colnames(c2df) <- c('Die Roll','Count')


hist1 <- plot_ly(c1df,x = c1df$`Die Roll`,y=c1df$Count,type='bar')%>%
  layout(title = '\nDistribution of 3d6 Rolls over 49 Rounds of Play', 
         xaxis = list(title = 'Die Results'), 
         yaxis = list(title = '# Occurrences'))
hist1 <- hist1 %>% layout(
  barmode="overlay",
  bargap=0.1)
hist1

hist2 <- plot_ly(c2df,x = c2df$`Die Roll`,y=c2df$Count,type='bar')%>%
  layout(title = '\nDistribution of 1d20 Rolls over 49 Rounds of Play', 
         xaxis = list(title = 'Die Results'), 
         yaxis = list(title = '# Occurrences'))
hist2 <- hist2 %>% layout(
  barmode="overlay",
  bargap=0.1)
hist2


```

# Histogram charts for larger number of sample rolls

```{r}

c3df <- as.data.frame(table(demo_3df['X3d6']))
colnames(c3df) <- c('DieRoll','Count')

c4df <- as.data.frame(table(demo_3df['X1d20']))
colnames(c4df) <- c('DieRoll','Count')

hist3 <- plot_ly(c3df,x=~DieRoll,y=~Count,type='bar')%>%
  layout(title = '\nDistribution of 3d6 Rolls over 10000 Samples', 
         xaxis = list(title = 'Summed Die Roll Results'), 
         yaxis = list(title = '# Occurrences'))
hist3 <- hist3 %>% layout(
  barmode="overlay",
  bargap=0.1)
hist3

hist4 <- plot_ly(c4df,x=~DieRoll,y=~Count,type='bar')%>%
  layout(title = '\nDistribution of 1d20 Rolls over 10,000 Samples', 
         xaxis = list(title = 'Single Die Roll Results'), 
         yaxis = list(title = '# Occurrences'))
hist4 <- hist4 %>% layout(
  barmode="overlay",
  bargap=0.1)
hist4

c5df <- merge(c4df, c3df, by='DieRoll',all.x=TRUE)
c5df[is.na(c5df)] <- 0 

plot5 <- plot_ly(c5df,x=~DieRoll,y=~c5df$Count.x,type='scatter',mode='lines+markers',name='1d20 = Stones')%>%
  layout(title = '\nComparison of Roll Distributions over 10,000 Samples', 
         xaxis = list(title = 'Die Roll Results'), 
         yaxis = list(title = '# Occurrences'))
plot5 <- plot5 %>% add_trace(y=~c5df$Count.y,name='3d6 = Bones')
plot5



```

# Area charts for game balances and flow

```{r}
demogamedata_1['SessRd'] <- do.call(paste,demogamedata_1[c('Session','Round')])
demogamedata_1['SessRd2'] <- demogamedata_1['Session']*100+demogamedata_1['Round']

gsfig <- plot_ly(demogamedata_1,x=~as.factor(SessRd2),y=~Treasury,type='scatter',mode='none', 
                 stackgroup='one',name='Treasury')%>%
  layout(title = '\nDemo 1', 
         xaxis = list(title = 'Round'), 
         yaxis = list(title = 'Coin Balance'),width=500)
gsfig <-gsfig %>% add_trace(y = ~P1_Bag, name = 'Player 1 Bag')
gsfig <-gsfig %>% add_trace(y = ~P2_Bag, name = 'Player 2 Bag')
gsfig <-gsfig %>% add_trace(y = ~P3_Bag, name = 'Player 3 Bag')
gsfig <-gsfig %>% add_trace(y = ~P4_Bag, name = 'Player 4 Bag')
gsfig

demogamedata_2['SessRd'] <- do.call(paste,demogamedata_2[c('Session','Round')])
demogamedata_2['SessRd2'] <- demogamedata_2['Session']*100+demogamedata_2['Round']

gsfig2 <- plot_ly(demogamedata_2,x=~as.factor(SessRd2),y=~Treasury,type='scatter',mode='none', 
                 stackgroup='one',name='Treasury')%>%
  layout(title = '\nDemo 2', 
         xaxis = list(title = 'Round'), 
         yaxis = list(title = 'Coin Balance'),width=500)
gsfig2 <-gsfig2 %>% add_trace(y = ~P1_Bag, name = 'Player 1 Bag')
gsfig2 <-gsfig2 %>% add_trace(y = ~P2_Bag, name = 'Player 2 Bag')
gsfig2 <-gsfig2 %>% add_trace(y = ~P3_Bag, name = 'Player 3 Bag')
gsfig2 <-gsfig2 %>% add_trace(y = ~P4_Bag, name = 'Player 4 Bag')
gsfig2

```

# Occurrence frequency charts for rules

```{r}
# subset dataframe to only the rules columns
rulecols <- c('P1','P2','P3','P4')
demorules_df1 <- as.data.frame(matrix(0,nrow=nrow(demogamedata_1),ncol=4))
colnames(demorules_df1) <- rulecols

for (eachcol in rulecols) {
  demorules_df1[,eachcol] <- demogamedata_1[,eachcol]
}

# make a pretty heatmap with pheatmap
rulefig1 <- pheatmap(t(demorules_df1),color=colorRampPalette(brewer.pal(n=8,name='Blues'))(100),
         cluster_rows = FALSE, cluster_cols = FALSE, show_rownames = TRUE,labels_col = unlist(demogamedata_1['SessRd2']),
         angle_col = 45, number_format = '%.0f', xlab = 'Round', ylab = 'Player',number_color='white',
         main='Demo 1: Rule Occurence by Round and Player',legend= TRUE,display_numbers=TRUE)
rulefig1

```

```{r}
demorules_df2 <- as.data.frame(matrix(0,nrow=nrow(demogamedata_2),ncol=4))
colnames(demorules_df2) <- rulecols

for (eachcol in rulecols) {
  demorules_df2[,eachcol] <- demogamedata_2[,eachcol]
}

rulefig2 <- pheatmap(t(demorules_df2),color=colorRampPalette(brewer.pal(n=8,name='Blues'))(100),
         cluster_rows = FALSE, cluster_cols = FALSE, show_rownames = TRUE,labels_col = unlist(demogamedata_2['SessRd2']),
         angle_col = 45, number_format = '%.0f', xlab = 'Round', ylab = 'Player',number_color='white',
         main='Demo 2: Rule Occurence by Round and Player',legend= TRUE,display_numbers=TRUE)
rulefig2
```


# Charts for Game length and Player win distribution

```{r}

hist6 <- plot_ly(samplerunstowin,x=~RunsToWin,y=~Occurrence,type='bar')%>%
  layout(title = '\nDistribution of Number of Rounds until Victory (1000 trials)', 
         xaxis = list(title = 'Number of Rounds'), 
         yaxis = list(title = '# Occurrences'))
hist6 <- hist6 %>% layout(
  barmode="overlay",
  bargap=0.1)
hist6

hist7 <- plot_ly(samplewinners,x=~Player,y=~Wins,type='bar')%>%
  layout(title = '\nDistribution of Wins by Player (1000 trials)', 
         xaxis = list(title = 'Player'), 
         yaxis = list(title = '# Wins'))
hist7 <- hist7 %>% layout(
  barmode="overlay",
  bargap=0.1)
hist7

#write.csv(samplerunstowin, file = "samplerunstowin.csv", fileEncoding = "UTF-8")

```

# Build the DOE matrix for possible game variations

```{r}
set.seed(222)

sbdoe_factors = list(Rule9Ante="",Rule13Steal="",Rule14Save="",ReEntry="",StartBag="",NPlayers="",
                                      StdBonus="",StartBonusDie="")

sbdoe_factors2 = list('Rule 9: Ante T/F','Rule 13: Steal T/F','Rule 14: Save T/F','Rule 17: ReEntry T/F',
                      'StartBag 20/40','NPlayers 3-4','StdBonus 5/10','StartBonusDie 1/3')
sbdoe_names = paste(rep('Trial',times=16),seq(1,16))

# run the DOE design using FrF2(), and store in a dataframe
stonesandbones_DOE <- FrF2(16,factor.names = sbdoe_factors)
sbdoe_mtx <- as.data.frame(matrix(as.integer(as.matrix(stonesandbones_DOE)),ncol=ncol(stonesandbones_DOE)))
colnames(sbdoe_mtx) <- sbdoe_factors2
sbdoe_mtx[sbdoe_mtx==-1] <- 0

# collect some basic statistics
nfeats = rowSums(sbdoe_mtx)
nshows = colSums(sbdoe_mtx)

# make a pretty heatmap with pheatmap
pheatmap(sbdoe_mtx,color=colorRampPalette(rev(brewer.pal(n=3,name='Blues')))(100),
         cluster_rows = FALSE, cluster_cols = FALSE, labels_col = sbdoe_factors2, labels_row = sbdoe_names, 
         angle_col = 315, number_format = '%.0f',
         main='Stones and Bones DOE: 16 trials, 8 features',legend= FALSE,display_numbers=TRUE)

sbdoe_factors3 = list('Rule9','Rule13','Rule14','Rule17',
                      'StartBag','NPlayers','StdBonus','StartBonusDie')
colnames(sbdoe_mtx) <- sbdoe_factors3
# Write out data to a CSV file 
write.csv(sbdoe_mtx, file = "DOE_matrix.csv", fileEncoding = "UTF-8")


```

# Import the DOE results, calculate metrics, store to a dataframe, make charts

```{r}

DOE_iteration <- 1
ntrials <- 16

doe_metrics_df <- as.data.frame(matrix(0,nrow=ntrials,ncol=6))
colnames(doe_metrics_df) <- c('Trial','MeanGameDur','VarGameDur','PctWon','NReversals','PlActive')

sbdoe_mtx <- as.data.frame(read.csv(file="DOE_matrix.csv",header=TRUE,sep=",",na.strings=c("")))

for (t in 1:ntrials) {
  
  trial = t-1
  # open the file
  doe_filename <- paste('TavernGameHist',1,'-',trial,'.csv',sep='')
  doe_trialdata <- as.data.frame(read.csv(file=doe_filename,header=TRUE,sep=",",na.strings=c("")))
  
  doe_metrics_df$Trial[t] <- trial
  
  # calculate metrics
  # mean game length
  doe_metrics_df$MeanGameDur[t] <- round(mean(doe_trialdata$RunsToWin),1)
  # stdev game length
  doe_metrics_df$VarGameDur[t] <- round(sd(doe_trialdata$RunsToWin),1)
  
  
  # % games won before round 50
  doe_metrics_df$PctWon[t] <- round(nrow(doe_trialdata[which(doe_trialdata$Winner!=''),])/10,1)
  
  # number of reversals per game (or session?)
  doe_metrics_df$NReversals[t] <- round(mean(doe_trialdata[,4]),1)
  
  # player active time
  # = average activerounds per player as % of game length
  # first check the number of active players, 4 or 3
  trialplayers <- 4 - sbdoe_mtx$NPlayers[t]
  doe_metrics_df$PlActive[t] <- round(mean(doe_trialdata$PlayerActiveRounds 
                                    / trialplayers / doe_trialdata$RunsToWin) * 100,1)
  
}

base_dur = doe_metrics_df$MeanGameDur[12]
base_var = doe_metrics_df$VarGameDur[12]

for (t in 1:ntrials) {
  
  # calculate the rho value based on the sample variances
  rho1 <- ((base_var + doe_metrics_df$VarGameDur[t])^2)
  rho2 <- ((base_var)^2)+((doe_metrics_df$VarGameDur[t])^2)
  rho <- (rho1/rho2)
  
  print(paste(rho1, rho2, rho))
  # look up the t-statistic based on the rho
  
  tstat <- qt(0.975,rho,lower.tail=TRUE)
  
  term1 = sqrt((base_var^2)+(doe_metrics_df$VarGameDur[t]^2))

  
  print(paste(t, rho, tstat))
  
}


# add 95% confidence interval data to the dataframe, as mean +/- 1.96*CI
doe_metrics_df['95ConfLower'] <- round(base_dur - doe_metrics_df$MeanGameDur[t] - tstat*term1,1)
doe_metrics_df['95ConfUpper'] <- round(base_dur - doe_metrics_df$MeanGameDur[t] + tstat*term1,1)


print(doe_metrics_df)

# Write out metricsdata to a CSV file 
write.csv(doe_metrics_df, file = "DOE_metrics.csv", fileEncoding = "UTF-8")


# make a scoring dataframe for trials
doe_scoring_df <- as.data.frame(matrix(0,nrow=ntrials,ncol=5))
colnames(doe_scoring_df) <- c('Trial','95ConfLower','PctWon','NReversals','PlActive')

doe_scoring_df$Trial <- doe_metrics_df$Trial
doe_scoring_df$'95ConfLower' <- (doe_metrics_df$'95ConfLower'>10)*1.0
doe_scoring_df$PctWon <- (doe_metrics_df$PctWon>80)*1.0
doe_scoring_df$NReversals <- (doe_metrics_df$NReversals>1.5)*1.0
doe_scoring_df$PlActive <- (doe_metrics_df$PlActive>75)*1.0
print(doe_scoring_df)

# make a summary scoring dataframe for trials
doe_sumscore_df <- as.data.frame(matrix(0,nrow=ntrials,ncol=2))
colnames(doe_sumscore_df) <- c('Trial','Score')
doe_sumscore_df$Trial <- doe_metrics_df$Trial
doe_sumscore_df$Score <- doe_scoring_df$`95ConfLower`+ doe_scoring_df$PctWon + doe_scoring_df$NReversals + doe_scoring_df$PlActive

# make a chart out of that dataframe
fig_doescore <- plot_ly(doe_sumscore_df,x=~seq(1,16),y=~Score,type='bar')%>%
  layout(title = '\nNumber of Metrics achieved by each DOE Trial', 
         xaxis = list(title = 'Trial'), 
         yaxis = list(title = 'Metrics Success Score'))
fig_doescore <- fig_doescore %>% layout(
  barmode="overlay",
  bargap=0.1)

fig_doescore

# then calculate the impacts for each of the 8 variables, by averaging the performance on each metric?
doe_impact_df <- as.data.frame(matrix(0,nrow=8,ncol=3))
colnames(doe_impact_df) <- list('Parameter','0','1')
doe_impact_df$Parameter <- sbdoe_factors2

##### Leaving the stub below as TBD for later maturation
for (f in sbdoe_factors2) {
  
  # subset metrics file with and without the chosen factor
  
  # calculate metrics for each subset
  
  # show chart comparing metrics for both states of the chosen factor
  
}


```

Calculate Approximate CI for each comparison pair.

```{r}

```


